### 2023-07-11 popl/main.tex
upto ex 3.8

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[utf8,acmsmall,review,screen,dvipsnames,anonymous]{acmart}

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{glossaries}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}

\setul{0.95ex}{0.3ex}

% https://tex.stackexchange.com/questions/648845/sans-serif-uppercase-greek-no-longer-showing-in-acmart
\DeclareMathAlphabet{\mathsf}{OT1}{LibertinusSans-LF}{m}{n}
\SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}

\DeclareMathAlphabet{\mathtt}{OT1}{lmtt}{m}{n}
\SetMathAlphabet{\mathtt}{bold}{OT1}{lmtt}{bx}{n}

\input{cmds.tex}

\loadglsentries{acronyms}
\makeglossaries

%% NOTE that a single column version may be required for 
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to 
%% \documentclass[manuscript,screen]{acmart}
%% 
%% To ensure 100% compatibility, please check the white list of
%% approved LaTeX packages to be used with the Master Article Template at
%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
%% before creating your document. The white list page provides 
%% information on how to submit additional LaTeX packages for 
%% review and adoption.
%% Fonts used in the template cannot be substituted; margin 
%% adjustments are not allowed.
%%
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
  2024}{London, UK}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03-05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Secure Composition of Robust and Optimising Compilers}
% Composing Secure and Optimising Compilers
%

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Matthis Kruse}
% \authornote{Both authors contributed equally to this research.}
\email{matthis.kruse@cispa.de}
\orcid{0000-0003-4062-9666}
\affiliation{%
  \institution{CISPA Helmholtz Center for Information Security and Saarland University}
  \streetaddress{Stuhlsatzenhaus 5}
  \city{Saarbr{\"u}cken}
  \state{Saarland}
  \country{Germany}
  \postcode{66123}
}

\author{Michael Backes}
\email{director@cispa.de}
%\orcid{0000-0002-7130-9211}
\affiliation{%
  \institution{CISPA Helmholtz Center for Information Security}
  \streetaddress{Stuhlsatzenhaus 5}
  \city{Saarbr{\"u}cken}
  \state{Saarland}
  \country{Germany}
  \postcode{66123}
}

\author{Marco Patrignani}
\orcid{0000-0003-3411-9678}
\email{marco.patrignani@unitn.it}
\affiliation{%
  \institution{University of Trento}
  \streetaddress{Via Sommarive, 9}
  \city{Povo}
  \country{Italy}
  \postcode{38123}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Kruse, Backes, and Patrignani}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}

  \MKin{context}
  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  \MKin{need}
  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness
  \MKin{task}
  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
  \MKin{object+findings}
  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
  \MKin{object}
  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  \MKin{conclusion}
  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  Some results of this work are implemented in the Coq proof assistant.

\begin{center}\small\it
	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}.
	Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
	a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
	and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
	}
\end{center}
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
  <concept>
  <concept_id>10002978.10002986.10002989</concept_id>
  <concept_desc>Security and privacy~Formal security models</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[500]{Security and privacy~Formal security models}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Memory-safety, Secure Compilation, Privacy}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction\pages{4}}\label{sec:introduction}

\gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.

Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
\begin{example}[strncpy]\label{ex:strncpy}
Consider the \texttt{C} program below that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}:

\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
void strncpy(size_t n, char *dst, char *src) {
  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
    dst[i] = src[i];
  }
}
\end{lstlisting}

For sake of simplicity, assume that the length of the array \texttt{dst} points to is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
Even under these assumptions, there is a subtle \gls{ms} vulnerability in the program: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}:
If \texttt{n} is at least as large as the length of the array \texttt{src}, the memory location past the last element will be read.
To prevent this vulnerability, one can use a compilation pass that enforces spatial memory safety, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.

Unfortunately, fixing \gls{ms} is not enough to have a secure program, because \texttt{strncpy} is susceptible to timing attacks.
This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
Thus, the program execution time is proportional to the length of the array that \texttt{src} points to.
% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
Also in this case there exist compiler passes that can rewrite such programs~\cite{cauligi2019fact} so that they are no longer vulnerable to timing attacks.

These fixes alone are not enough to make this program secure in a robust setting, i.e., a malicious context could trigger a violation of either \gls{ms} or \gls{cct}.
Consider a program context that calls \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
This leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the non-nullability guarantees of the source language.
For compilers to target languages that provide capabilities (such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}), which are unforgable handles that encapsulate information about a pointer, they can leavarage the information stored in these capablities to ensure the non-nullability guarantees in the target.
\end{example}

This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and \gls{cct}.
Then, this paper defines compiler composition and formalises what kind of security properties does the composition of two compiler passes preserve, given what the individual passes preserve.
The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.

In summary, this paper makes the following contributions:
\begin{itemize}
  \item %
        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
        % 
        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation.
        % 
        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
        % 
        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}.
%        \MPin{
%        	conclude that did this prop in a single pass, we show how to do it compositionally
%        }

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
        For example, a compiler that is secure with respect to \gls{tms} can be composed with a compiler that is secure with respect to \gls{sms}, resulting in a compiler that is secure with respect to \gls{ms}.
%       \MPin{
%      		find way to connect to previous contrib
%       }
        % 
        Finally, this paper provides an important corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This corollary is crucial for reordering compiler optimisation passes as well as secure compilation passes and thus generating secure and efficient code.

  \item %
        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
        To this end, it presents a compilation chain consisting of several passes.
        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and a stricter variant of \gls{cct}, respectively.
        Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
        The formalisation of this case study showcases the power of the presented framework: Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.

  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
\end{itemize}

This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).

\contribution{Open Source \& Technical Report} A technical report with the omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
% \MKin{cite/link}

\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}

To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).

\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}

This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security-relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
}

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

\begin{definition}[Property Satisfaction]\label{def:propsat}
  \bul{$\sat{\progvar}{\pi}$}
  $\isdef$
  if \iul{$\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
  then \oul{$\trace\in\pi$}.
\end{definition}

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

\begin{definition}[Robust Satisfaction]\label{def:proprsat}
  \bul{$\rsat{\progvar}{\pi}$}
  $\isdef$ \iul{$\forall \contextvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
\end{definition}

\begin{example}[Double Free in Bluetooth Subsystem]
  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
\end{example}

\subsection{Secure Compilers}\label{subsec:bg:rtp}

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.

\begin{definition}[Robust Compilation]\label{def:rtp}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
  %, iff 
  $\isdef$
  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\Ltms\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
\end{definition}

Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.

\begin{example}[Types]
  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
\end{example}

\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.

\subsection{Specification Trace Model}\label{subsec:propdefs}

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone are enough to provide a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

\subsubsection{Temporal Memory Safety}

\gls{tms} is a safety property that describes that an unallocated object must not be used in any way.
Moreover, the property requires that all allocated objects must be deallocated at some point.

\begin{definition}[\glsfirst{tms}]\label{def:trace:tmsdef}
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

\subsubsection{Spatial Memory Safety}

\gls{sms} prohibits out-of-bounds accesses:

\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
  $
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
  $
\end{definition}

\subsubsection{Memory Safety}

Full \gls{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}

\subsubsection{Strict Cryptographic Constant Time}

\gls{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution. not as easy to check with just monitors.
To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace (inspired by earlier work~\cite{almeida2017jasmin}).

\begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
  $
  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                 \end{array}\right.\right\}
  $
\end{definition}

\subsubsection{Memory Safe, Strict Cryptographic Constant Time}

The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
Since \gls{ms} and \gls{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls{ms} and \gls{scct}.

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\mssafe\cap\scctsafe
  $
\end{definition}

%\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
\subsection{Monitors\pages{1.5}}\label{subsec:monitors}

\begin{figure}[h!]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}

Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

\begin{definition}[Monitor Satisfaction]\label{def:monsat}
  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
\end{definition}

\subsubsection{Monitor for \glsfirst{tms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
Contrary to other monitors in this paper, the multi-step relation of the \gls{tms} monitor is non-standard:
\begin{center}
  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
  %
  \typerule{tms-refl}{
    \tmsmonitor.\text{allocated}=\emptyset
  }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
  }{tms-refl}
  %
  \typerule{tms-ign-trans}{
    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
  }{tms-ign-trans}
  %
  \typerule{tms-trans}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}
    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
  }{tms-trans}
\end{center}
\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that no more locations should be allocated.
This rejects the behavior of programs that forget to free memory.

\begin{center}
  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{tms}.''}$\;$\\
  %
  \typerule{tms-alloc-authentic}{
  }{
    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc-authentic}
  %
% \typerule{tms-dealloc-authentic}{
% }{
%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
% }{tms-dealloc-authentic}
  %
% \typerule{tms-use-authentic}{
% }{
%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
% }{tms-use-authentic}
  %
  \typerule{tms-branch-authentic}{
  }{
    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
  }{tms-branch-authentic}
  %
% \typerule{tms-binop-authentic}{
% }{
%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
% }{tms-binop-authentic}
% %
% \typerule{tms-empty-authentic}{
% }{
%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
% }{tms-empty-authentic}
% %
  \typerule{tms-abort-authentic}{
  }{
    \tmstraceagree[]{\lightning}{\absterm}
  }{tms-abort-authentic}
\end{center}

The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.

\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
\end{lemma}

\begin{example}[A program not satisfying \gls{tms}]
  Consider the following \texttt{C++11} program that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.

  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
int main() {  // allocates 12 chars containing a greeting message
  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$

  strncpy(12, to, greetings);
  delete to;

  printf("%cOPL\n", to[6]);
}
  \end{lstlisting}
  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), it is easy to see that it does not deallocate its arguments.
  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
  However, having returned from \texttt{strncpy}, the \texttt{main} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  It is easy to see that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
\end{example}

\subsubsection{Monitor for \glsfirst{sms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{sms-uninteresting}{
% }{
%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% }{sms-uninteresting}
% %
% \typerule{sms-abort}{
% }{
%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% }{sms-abort}
  %
  \typerule{sms-use}{
    (\loc;\sandboxtag;m)\in\smsmonitor &
    n<m
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
  }{sms-use}
  %
  \typerule{sms-alloc}{
    (\loc;\sandboxtag;m)\notin\smsmonitor
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
  }{sms-alloc}
\end{center}
The state of the monitor for \gls{sms} is a set containing tuples of locations, control-tags, and the allocation size.
In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
The trace agreement is entirely straightforward and similar to the one for \gls{tms}.
% \begin{center}
%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
%   %
%   \typerule{sms-alloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
%   }{sms-alloc-authentic}
%   %
%   \typerule{sms-dealloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
%   }{sms-dealloc-authentic}
%   %
%   \typerule{sms-use-authentic}{
%   }{
%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
%   }{sms-use-authentic}
%   %
%   \typerule{sms-branch-authentic}{
%   }{
%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{sms-branch-authentic}
%   %
%   \typerule{sms-binop-authentic}{
%   }{
%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
%   }{sms-binop-authentic}
%   %
%   \typerule{sms-empty-authentic}{
%   }{
%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
%   }{sms-empty-authentic}
%   %
%   \typerule{sms-abort-authentic}{
%   }{
%     \smstraceagree[]{\lightning}{\absterm}
%   }{sms-abort-authentic}
% \end{center}

\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
\end{lemma}

\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
  For the sake of this example, the pointers have been allocated by the component and passed to the context.
  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
  Because of this, there is an out of bounds\MP{hyphenate: out-of-bounds} memory access right before exiting the function.
  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.

  It is easy to see that,\MP{erase. we don't know what's easy} after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
  Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}. Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.
\end{example}

\subsubsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}

\begin{center}
  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
  \typerule{ms-step}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
  }{
    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
  }{ms-step}
\end{center}
The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep. The trace agreement similarly just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
% \begin{center}
%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
%   %
%   \typerule{ms-authentic}{
%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
%   }{
%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
%   }{ms-authentic}
% \end{center}

\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{lem:mon:msafe}
  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
\end{lemma}

\subsubsection{Monitor for \glsfirst{scct}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
  %
  \typerule{scct-none}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
  }{scct-none}
  %
  \typerule{scct-abort}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
  }{scct-abort}
\end{center}
The monitor state for the \gls{scct} monitor is completely empty, since it does not need to keep track of information.
As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
\begin{center}
  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{cct}.''}$\;$\\
  %
  \typerule{scct-low-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
  }{scct-low-authentic}
  %
  \typerule{scct-high-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
  }{scct-high-authentic}
  %
  \typerule{scct-empty-authentic}{
  }{
    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
  }{scct-empty-authentic}
  %
  \typerule{scct-abort-authentic}{
  }{
    \sccttraceagree[]{\lightning}{\absterm}
  }{scct-abort-authentic}
\end{center}
Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.

\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
\end{lemma}

\begin{example}[Data-independent timing mode]
  Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
  The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
  In terms of the abstract monitor trace, this is just $\absev{Any}$.
  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
\end{example}

\subsubsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}

The combination of monitors for \gls{ms} and \gls{scct} yields one for \gls{msscct}.
The construction is entirely similar to the one for \gls{ms} (\Cref{subsubsec:ms}).

\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{lem:mon:msscctsafe}
  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
\end{lemma}

\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}

This section presents the key meta-theoretic results of this paper concerning sequential compiler composition (and of optimisation passes) (\Cref{sec:sequential}) and concerning other kinds of compiler composition (\Cref{sec:other-compos}).
% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all? may reveal identity even though its double-blind}. % MP: not needed at all, ever, workshop work does not need to be cited in the main work

\subsection{Secure Sequential Composition}\label{sec:sequential}

The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
\MP{?canetti?} %% protocols under certain security notions are not secure when composed
The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
\MPin{
	section 3.1 never defined the $\cap$ operator on properties, nor on classes.
	it uses the 'and'.
	rectify
}

\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
\end{theorem}

Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.

\subsubsection{Securing Optimisations}
Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
% \MPin{survey to cite? this is a bit informal MK: hard to find a survey on different compiler passes... I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
% 
% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
% That way, an order of optimisations can be found by hand that yields the best results.
% \MP{what's the point of the previous 3 sentences?(commented)}
Consider the following two LLVM optimisation passes: \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
The order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}. \Coqed
\end{corollary}

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
Examples of this are present in industry:
Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
Both languages can be used at the same time in one project~\cite{androidstudio}.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
The definition of upper composition is as follows: given a program $\mathtt{p}$, its compiled $\trg{L}$ counterpart is obtained by plugging $\mathtt{p}$ into $\cc{\src{L}}{\trg{L}}$ if $\mathtt{p}\in\src{L}$ or by plugging $\mathtt{p}$ into $\cc{\obj{L}}{\trg{L}}$ if $\mathtt{p}\in\obj{L}$, written as $\cc{\src{L}+\obj{L}}{\trg{L}}$.
\MP{
	put into def env
}
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
\end{theorem}

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
The %proposed
definition of lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ is as follows: given a program $\src{p}$, its compiled $\trg{L}$ or $\obj{L}$ program is obtained by performing an internal decision and threading $\src{p}$ into $\cc{\src{L}}{\trg{L}}$ or into $\cc{\src{L}}{\obj{L}}$ depending on that decision.
\MP{into def}
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
\end{theorem}

%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.

%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

This paper considers \gls{cct} as an architecture-dependent property writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
In spirit of this, language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
  &
  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
  &
  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
  \end{aligned}
\end{gather*}
% \MPin{
%   this latex had many issues.
%   the \& were at different spots
%   it used hspace to space things instead of \&
%   and it did not use \& to align the following things on a line
%   and it did not use multiple aligned into the gather
%   (what's the point of gather if you don't have several aligned inside? use a single align* instead)
% }
% oh... that's good to know! the tech report is also riddled with this... :/

Above is a fraction of the shared syntax of all the programming languages of this paper.
\MPin{why a fraction? what exprs are missing
MK: Things that are missing here are: Evaluation contexts, states $\Omega$ and transitively other definitions affected by this
}
The syntax and semantics are mostly standard and therefore omitted.
However, when allocating arrays with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the cells and $\expr[_{2}]$ the overall size of an array.

\MPin{
	below is semantics info. should go with their symbols and the semantics rules.
	and a better description of the semantics symbols.
  MK: this moved to appendix! should we bring it back? what exactly should be left out?
}
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control-tag $\sandboxtag$ (either $\ctx$ or $\comp$) to know the heap they point into.
Other metadata in $\Omega$ entails a list of functions with information on whether they belong to the context or the component.
That way, when calling a function, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, functions are second-class.
So, the only values are pairs, pointers, and natural numbers.
Another non-standard part of the languages is the concept of the poison-tag.
The poison-tag marks an array as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that a program like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ yields a trace $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$ when executing.

\begin{gather*}
  \begin{aligned}
  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
  \end{aligned}
\end{gather*}

Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.

\begin{center}
  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
  %
  \typerule{$e-\text{ret}$}{
    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-ret-notsame}
  %
  \typerule{$e-\text{call}-\text{notsame}$}{
    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-call-notsame}
\end{center}

The environmental semantics is also mostly straightforward.
In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.

The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.

\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}

$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.\MP{fabulous paper from amal?relevant?}
\MP{and therefore omitted?}
The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
\MPin{ok let's add details}

\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{thm:wt:tms}
  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
\end{theorem}

\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}

$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
%
The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
\MP{ok, details}

\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$.
The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.

\subsection{$\Lscct$: A Language with a Data Independent Timing Mode}\label{subsec:lscct}

While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
For the other languages seen earlier, this mode is intuitively always-on.
\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
If the register is turned on, no secrets can appear on traces.
If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
\Cref{ex:lscct} illustrates these changes.
\MPin{formalisation additions in algin* env}

The projection to the specification trace model is mostly straightforward.
But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
With this technical setup, the information wether\MP{spelling} a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.

\begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
  Consider the following $\Lscct$ program and the execution trace $\obj{\trace}$ juxtaposed with it.

  $$
  \begin{array}{lcc}
  \obj{\lwrdoit{0};} & \mid & \obj{\emptyevent}\\
  \obj{\llet{x_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock} \\
  \obj{\lset{x}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\ev{Branch}\ 0;\sandboxtag;\lock\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock} \\
  \obj{\lwrdoit{1};} & \mid & \obj{\emptyevent}\\
  \obj{\llet{y_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock}\\
  \obj{\lset{y_{\lock}}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\emptyevent\cdot\widehat{Set}\ \loc_{y}\ 2;\sandboxtag;\lock}\\
  \end{array}
  $$

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event does not fire anymore.
  Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
  \MPin{shouldn't we have 2 traces, one with the bit set, one without}
  Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
  $\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock\cdot\ev{Branch\ 0}\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock\cdot\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock\cdot\ev{Set}\ \loc_{y}\ 2;\sandboxtag;\unlock$.
  Clearly, secrets are hidden ($\unlock$) after this projection for the section of the program where the data independent timing mode is turned on, while still containing the appropriate events to check for \gls{ms}.
\end{example}

\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
\begin{figure}[!h]
  \begin{center}
    \begin{tikzpicture}
      \node (S) {$\src{L_{\tmssafe}}$};
      \node[right=1.5 of S] (T) {$\trg{L}$};
      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};

      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);

      % Sections
      \node[above=1.0 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
      \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
      \node[right=1.75 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
      \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};

      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls{ms}+\gls{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);

      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure}
The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
\MPin{some text should introduce the compiler defs}

  \begin{align*}
%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                \left[\cca(\src{\expr})\right] %
                                                                                                 }{\labort[\trg]}}
  \end{align*}

The first compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
By inserting the check, the compiler ensures that execution does not proceed in such cases.
The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).

Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.

\begin{theorem}[Compiler $\cca$ is secure with respect to \gls{tms}]\label{thm:cca:rtp:tms}
  $\rtp{\cca}{\tmssafe}$ % \Coqed
\end{theorem}

\subsubsection{Proving Robust Safety Property Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for the following theorems of this section.
\MP{so what?}

When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
\MP{passive + not justified}
The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
\MP{no, the goal is to BT}
To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.\MP{passive}
By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.\MP{passive passive passive, and imprecise}
The proof now needs to relate the two execution with each other.
\Cref{fig:proofdiag:rtp} visualises this situation\MP{imprec}.
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
      % relative horizontal/vertical distance between states
      \pgfmathsetmacro{\hdist}{0.95}
      \pgfmathsetmacro{\vdist}{1.35}
      \pgfmathsetmacro{\halfvdist}{0.725}

      % row of src states
      \node[state] (srcempty) {$\src{\emptyset}$};
      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
        \node[state,right=\hdist of src\cur] (src\s) {$\src{\Omega_{\s}}$};
      }
      % row of trg states
      \node[state,below=\vdist of srcempty] (trgempty) {$\trg{\emptyset}$};
      \foreach \s in {1,w_1,p,w_2,2} {
        \node[state,below=\vdist of src\s] (trg\s) {$\trg{\Omega_{\s}}$};
      }
      %% illustrations
        % backtrans wrapper 1
        \draw[thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
        % backtrans wrapper 2
        \draw[thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
        % compiler correctness
        \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
      % state relations
      \path (srcempty) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
        (src1) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
        (srcp) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
        (src2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
        (srcw\string_1) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
        (srcw\string_2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
        % diagonals
        (srcw\string_1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
        (src1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
        (srcp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
        (trgp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
        ;
      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
      %\path[->,draw] (wrapper) -- (srcw\string_1);
      %\path[->,draw] (wrapper) -- (srcp);
      % steps
      \path[color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
        ;
      \path[color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
        (trgp) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
        ;
      % legend
      \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
        };
      \draw[thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
      \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
    \end{tikzpicture}
    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs.}\label{fig:proofdiag:rtp}
  \end{center}
\end{figure}
\MPin{
	thicker lines and dots needed, not visible otherwise
}

At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again. Recursive calls between context and component functions are disregarded.}.
Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmssafe}.

\MPin{
	why can we use a BT?
	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
	all these intuitions are missing.
	they are technical, so we only keep them here.

  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
}

\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}

\begin{center}
  $$
  \begin{array}{rcl}
    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
  \end{array}
  $$
\end{center}

The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
\MPin{
	show the result of the compilation as listing, and then describe the trace a bit more in detail ?
}

\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls{sms}]\label{thm:ccb:rtp:sms}
  $\rtp{\ccb}{\smssafe}$ % \Coqed
\end{theorem}

\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls{ms}]\label{thm:ccab:rtp:ms}
  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
\end{theorem}

\subsection{Optimising Compilers}\label{subsec:cs:opts}

\begin{center}
  $$
  \begin{array}{rcl}
    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{1}]}) \\
    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{2}]}) \\
  \end{array}
  $$
\end{center}

\begin{center}
  $$
  \begin{array}{rcll}
    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
  \end{array}
  $$
\end{center}
\MPin{
	substitution are a runtime semantics thing.
	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
  %
  \MKin{
  MK: No? I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics. I see how these two are different. However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to. If we don't do this, we essentially re-encode the definition of subst into the compiler
  }
}

The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
Both are secure with respect to \gls{ms}, the proof of which is relatively easy, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.

\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
  $\rtp{\ccdce}{\mssafe}$ \Coqed
\end{theorem}
\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
  $\rtp{\cccf}{\mssafe}$ \Coqed
\end{theorem}

With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:

\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls{ms}]\label{thm:cccfccdce:rtp:ms}
  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
\end{theorem}

\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}

\begin{center}
  $$
  \begin{array}{rcl}
    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})}
  \end{array}
  $$
\end{center}

Given the fact that $\Lscct$ provides a \gls{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls{cct}-mode.
The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
Because of this, the compiler is secure with respect to \gls{scct}, similarly proven as in \Cref{subsec:cs:tms}.

\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
\end{theorem}

\subsection{Robust Preservation of Intersection of Memory Safety and Strict Cryptographic Constant Time}

Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
%
	\MKin{ 
	Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}

	\MPin{yes it should}}
}

\section{Related Work\pages{2}}\label{sec:relwork}

This section compares existing work with the one presented in this paper.
To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey}.
This work has made use of that previous work.
While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.

\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}

Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.

\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}

The approach to cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.

\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}

This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
It is also interesting to extend our case-study and verify all results of it in Coq.
This way, it is possible to obtain an executable, secure compiler.
However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
% \begin{acks}
%   We would like to thank the anonymous reviewers for their feedback.
% \end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.