### 2024-05-27 csf2/acronyms.tex
nits of spacing + intro

\newacronym{rtp}{RTP}{Robust Trace-Property Preservation}
\newacronym{tms}{TMS}{Temporal Memory Safety}
\newacronym{sms}{SMS}{Spatial Memory Safety}
\newacronym{ms}{MS}{Memory Safety}
\newacronym{specms}{SpecMS}{Speculation Memory Safety}
\newacronym{cct}{CCT}{Cryptographic Constant Time}
\newacronym{scct}{sCCT}{Strict Cryptographic Constant Time}
\newacronym{ss}{SS}{Speculative Safety}
% \newacronym{msscct}{MS+sCCT}{Memory Safety and Strict Cryptographic Constant Time}
% \newacronym{mssscct}{MS+sCCT+$\mathghost$}{Memory Safety and Strict Cryptographic Constant Time with Speculation}
\newacronym{dce}{DCE}{Dead Code Elimination}
\newacronym{cf}{CF}{Constant Folding}

\newacronym{msscct}{MCT}{Memory Safety and Strict Cryptographic Constant Time}
\newacronym{mssscct}{MCT$\mathghost$}{Memory Safety and Strict Cryptographic Constant Time with Speculation}