### 2024-10-14 tech/acronyms.tex
tech

\newacronym{rtp}{RTP}{Robust Trace-Property Preservation}
\newacronym{tms}{TMS}{Temporal Memory Safety}
\newacronym{sms}{SMS}{Spatial Memory Safety}
\newacronym{ms}{MS}{Memory Safety}
\newacronym{cct}{CCT}{Cryptographic Constant Time}
\newacronym{scct}{sCT}{Strict Cryptographic Constant Time}
\newacronym{msscct}{MS+sCCT}{Memory Safety and Strict Cryptographic Constant Time}
\newacronym{dce}{DCE}{Dead Code Elimination}
\newacronym{cf}{CF}{Constant Folding}
\newacronym{ss}{SS}{Speculation Safety}
\newacronym{specms}{SpecMS}{Speculation Memory Safety}