### 2024-10-15 tech/tex/monitors.tex

\subsection{Monitors}\label{subsec-cs-monitors}

This part presents monitors for the properties presented in \Cref{subsec-cs-properties}.
Since all properties are safety properties, it is enough to consider trace-prefixes, i.e., finite lists of events.

In this section, we use the syntax $\exists\noncrashMonitor{\varMonitor}, \phi$ as shorthand for $\exists\varMonitor, \varMonitor\not=\lightning \text{ and } \phi$, where $\phi$ is some proposition that may depend on $\varMonitor$.
Essentially, whenever $\noncrashMonitor{\varMonitor}$ occurs, there is a constraint that $\varMonitor$ must not be $\lightning$.

\input{tex/monitors/tmsmon}
\input{tex/monitors/smsmon}
\input{tex/monitors/dual}
\input{tex/monitors/msmon}
\input{tex/monitors/scctmon}
\input{tex/monitors/msscctmon}
\input{tex/monitors/ssmon}
\input{tex/monitors/specmsmon}