### 2024-10-15 tech/tex/monitors/scctmon.tex

\subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}

\begin{definition}{\definitionlabel[\gls*{scct} Monitor State]{monitor-scct}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract State)}~\scctMonitor &\isdef \emptyset \mid \lightning\\
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[\gls*{scct} Monitor Reductions]{monitor-scct-steps}}
  \begin{center}
    \judgbox{\monitorcheck[]{\scctMonitor}{\scctMonitor[']}{\scctEvent}}{,,Given abstract state $\scctMonitor$, step to abstract state $\scctMonitor'$\\ emitting an abstract event $\scctEvent$.''}

    \typerule{scctmon-empty}{}{
      \monitorcheck[]{\scctMonitor}{\scctMonitor}{\emptyevent}
    }{scctmon-empty}
    \typerule{scctmon-any}{
    }{
      \monitorcheck[]{\scctMonitor}{\scctMonitor}{\scctAny[\unlock]}
    }{scctmon-any}
    \typerule{scctmon-crash}{}{
      \monitorcheck[]{\scctMonitor}{\lightning}{\lightning}
    }{scctmon-crash}
  \end{center}
  \begin{center}
    \judgbox{\monitorcheck{\scctMonitor}{\scctMonitor[']}{\scctEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\scctTrace$ is a list of events.''}

    \typerule{scctmon-refl}{
    }{
      \monitorcheck{\scctMonitor}{\scctMonitor}{\hole{\cdot}}
    }{scctmon-refl}
    %
    \typerule{scctmon-trans}{
      \monitorcheck[]{\scctMonitor}{\scctMonitor[']}{\scctEvent} \rulesep
      \monitorcheck{\scctMonitor[']}{\scctMonitor['']}{\scctTrace}
    }{
      \monitorcheck{\scctMonitor}{\scctMonitor['']}{\scctEvent\cdot\scctTrace}
    }{scctmon-trans}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[\gls*{scct}-Satisfaction via Monitor]{trace-is-scct-via-mon}}
  \begin{align*}
    \scctsafe\vdash\scctTrace &\isdef \exists\noncrashMonitor{\scctMonitor},\monitorcheck{\emptyset}{\scctMonitor}{\scctTrace}
  \end{align*}
\end{definition}