### 2024-10-15 tech/tex/monitors/specmsmon.tex

\subsubsection{\gls*{specms} Monitor}\label{subsubsec-mon-specms}
Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-msscct,subsubsec-mon-spec}. 
We refer to the resulting monitor state as $\specmsMonitor$.

\begin{definition}{\definitionlabel[\gls*{msscct}-Satisfaction via Monitor]{trace-is-msscct-via-mon}}
  \begin{align*}
    \msscctsafe\vdash\varTrace[\mssafe;\sssafe] &\isdef \exists\noncrashMonitor{\msscctMonitor},\monitorcheck{\emptyset}{\msscctMonitor}{\varTrace[\mssafe;\sssafe]}
  \end{align*}
\end{definition}