### 2024-10-15 tech/tex/monitors/ssmon.tex

\subsubsection{\gls*{ss} Monitor}\label{subsubsec-mon-spec}

\begin{definition}{\definitionlabel[\gls*{ss} Monitor State]{monitor-spec}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract State)}~\specMonitor &\isdef \emptyset \mid \lightning\\
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[\gls*{ss} Monitor Reductions]{monitor-spec-steps}}
  \begin{center}
    \judgbox{\monitorcheck[]{\specMonitor}{\specMonitor[']}{\specEvent}}{,,Given abstract state $\specMonitor$, step to abstract state $\specMonitor'$\\ emitting an abstract event $\specEvent$.''}

    \typerule{specmon-empty}{}{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\emptyevent}
    }{specmon-empty}
    \typerule{specmon-any-public}{
    }{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\specAny[\unlock]}
    }{specmon-any-public}
    \typerule{specmon-any-private}{
    }{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\specAny[\lock_{\text{NONE}}]}
    }{specmon-any-private}
    \typerule{specmon-crash}{}{
      \monitorcheck[]{\specMonitor}{\lightning}{\lightning}
    }{specmon-crash}
  \end{center}
  \begin{center}
    \judgbox{\monitorcheck{\specMonitor}{\specMonitor[']}{\specEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\specTrace$ is a list of events.''}

    \typerule{specmon-refl}{
    }{
      \monitorcheck{\specMonitor}{\specMonitor}{\hole{\cdot}}
    }{specmon-refl}
    %
    \typerule{specmon-trans}{
      \monitorcheck[]{\specMonitor}{\specMonitor[']}{\specEvent} \rulesep
      \monitorcheck{\specMonitor[']}{\specMonitor['']}{\specTrace}
    }{
      \monitorcheck{\specMonitor}{\specMonitor['']}{\specEvent\cdot\specTrace}
    }{specmon-trans}
  \end{center}
\end{definition}
Note that \trref{specmon-any-private} allows secrets to be exposed on the trace if the leak is not because of speculation.

\begin{definition}{\definitionlabel[\gls*{ss}-Satisfaction via Monitor]{trace-is-spec-via-mon}}
  \begin{align*}
    \sssafe\vdash\specTrace &\isdef \exists\noncrashMonitor{\specMonitor},\monitorcheck{\emptyset}{\specMonitor}{\specTrace}
  \end{align*}
\end{definition}