### 2024-10-22 spectre/acronyms.tex
fill parts of spectre table

\newacronym{rtp}{RTP}{Robust Trace-Property Preservation}
\newacronym{tms}{TMS}{Temporal Memory Safety}
\newacronym{sms}{SMS}{Spatial Memory Safety}
\newacronym{ms}{MS}{Memory Safety}
\newacronym{cct}{CCT}{Cryptographic Constant Time}
\newacronym{scct}{sCT}{Strict Cryptographic Constant Time}
\newacronym{msscct}{MS+sCCT}{Memory Safety and Strict Cryptographic Constant Time}
\newacronym{dce}{DCE}{Dead Code Elimination}
\newacronym{cf}{CF}{Constant Folding}
\newacronym{ss}{SS}{Speculation Safety}
\newacronym{specms}{SpecMS}{Speculation Memory Safety}

\newacronym{s}{S}{Software}
\newacronym{h}{H}{Hardware}
\newacronym{hs}{HS}{Hardware-Simulation}

\newacronym{seq}{SEQ}{Sequential, in-order execution of instructions.}
\newacronym{spec}{SPEC}{Speculative, in-order execution of instructions.}
\newacronym{seq-ooo}{SEQOOO}{Sequential, out-of-order execution of instructions.}
\newacronym{spec-ooo}{SPECOOO}{Speculative, out-of-order execution of instructions.}

\newacronym{memobs}{MEMobs}{Sees loads and store memory addresses.}
\newacronym{ctobs}{CTobs}{Sees loads, stores, and program counter memory addresses.}
\newacronym{pathctobs}{PATH-CTobs}{Constant-Time observer. Sees loads, stores, and program counter memory addresses, as well as all possible (speculative) paths.}
\newacronym{ct+pcobs}{CT+PCobs}{\acrshort{ctobs}-obsverver, but in speculative-mode only sees program counter.}
\newacronym{archobs}{ARCHobs}{Architectural observer. Sees any microarchitectural covert channel.}

\newacronym{ccseqct}{$\ccseqct$}{\acrshort{seq}-execution; \acrshort{ctobs}-observations.}
\newacronym{ccspecmem}{$\ccspecmem$}{\acrshort{spec}-execution; \acrshort{memobs}-observations.}
\newacronym{ccspecct}{$\ccspecct$}{\acrshort{spec}-execution; \acrshort{ctobs}-observations.}
\newacronym{ccspecpathct}{$\ccspecpathct$}{\acrshort{spec}-execution; \acrshort{pathctobs}-observations}
\newacronym{ccseqarch}{$\ccseqarch$}{\acrshort{seq}-execution; \acrshort{archobs}-observations}
\newacronym{ccspecarch}{$\ccspecarch$}{\acrshort{spec}-execution; \acrshort{archobs}-observations}
\newacronym{ccseqspecctpc}{$\ccseqspecctpc$}{\acrshort{seq}- and \acrshort{spec}-execution; \acrshort{ct+pcobs} observer}

\newacronym{fi}{FI}{Fence Insertion}
\newacronym{cmr}{CMR}{Cancelable Memory Requests}
\newacronym{dtt}{DTT}{Dynamic Taint Tracking}
\newacronym{msr}{MSR}{Model-Specific Register}
\newacronym{iso}{ISO}{Process Isolation}
\newacronym{rz}{RZ}{Register Zeroing}
\newacronym{sst}{SSt}{Shadow Stack}
\newacronym{ri}{RI}{Register Interlocking}
\newacronym{slh}{SLH}{Speculative Load Hardening}