### 2025-02-03 icfp/sections/intro.tex
sketch of intro

\section{Introduction}\label{sec:introduction}

Programming languages provide security enforcements, e.g., Rust memory safety by means of ownership semantics.

Compilers may fail to preserve these enforcements, even when proven correct.

This is why secure compilers have become an interesting avenue for research, but they are difficult to prove.

Compositionality to the rescue, proofs should compose to make it easier.

% a secure compiler, by definition, is supports source-independent linking
% vertical composition is a consequence of existence of backtranslations

- Demonstrate compositionality of secure compilation.

- Mechanization of core theory.

- Case Study demonstrating composition.