### 2025-02-05 icfp/sections/bg/hprop.tex

\subsection{Hyperproperties and (Robust) Satisfaction}

Hyperproperties~\cite{clarkson2008hyper} are strictly more expressive than ordinary properties. 
While a property is a set of traces, a hyperproperty is a set of properties. 
We can lift properties to hyperproperties by applying the powerset.
Hyperproperty satisfaction is defined as follows.

It turns out that classes of hyperproperties, that is, sets of hyperproperties, are equally expressive than simply hyperproperties themselves~\cite{clarkson2008hyper}.

% and so on